### 2023-04-17 manuscript.Rmd
changes made in response to suggestions from reviewers round 1

---
title: "Paleoenvironmental models for Australia and the impact of aridification on blindsnake diversification"
date: "`r Sys.Date()`"
author: ""
output: 
  # bookdown::html_document2:
  #   toc: true
  #   number_sections: FALSE
  bookdown::word_document2:
    reference_docx: [C:/Users/ST/Documents/repo/AusARG_blindsnake/custom-reference-doc.docx]
    number_sections: false
    fig_caption: yes
bibliography: [C:/Users/ST/Documents/repo/anu_thesis_texts/zotero_library.bib]
# biblio-style: "apalike"
csl: jbiogeography.csl
link-citations: true
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)
# knitr::opts_knit$set(root.dir = normalizePath("..")) #use grandparent folder working directory
# for including graphics we need full path with knitr
WD <- getwd()
```

```{r packages}
library(dplyr); library(readr); library(janitor); 
library(ape); library(phytools);
library(kableExtra); library(RPANDA)
```

**Running title** Decrease in blindsnake speciation correlate with aridification

**Authors**: Sarin Tiatragul^1^; Alexander Skeels^2,3^; J. Scott Keogh^1^

^1^Division of Ecology & Evolution, Research School of Biology, The Australian National University, Canberra 2601, ACT, Australia

^2^Department of Environmental Systems Sciences, Landscape Ecology, Institute of Terrestrial Ecosystems, ETH Zürich, Zurich 8092, Switzerland

^3^Swiss Federal Institute for Forest, Snow and Landscape Research WSL, Birmensdorf 8903, Switzerland

**Corresponding author**: Sarin Tiatragul (sarin.tiatragul@anu.edu.au)

# Acknowledgements
ST would like to thank J Rolland and F. Condamine for helpful advice about fitting diversification models with RPANDA. Thanks to NJ Matzke and O Jiménez-Robles for help with BioGeoBEARS. Special thanks to IG Brennan, D Esquerré, M Pepper, CJ Pavón-Vázquez, and O Hagen for helpful advice on the analysis and discussion. JSK thanks the Australian Research Council for ongoing support.

\newpage

# Abstracts and keywords

**Aim**: Shifts in diversification rates of Australian flora and fauna have been associated with aridification, but the relationship between diversification rates and aridity has never been quantified. We employed multiple approaches to accurately reconstruct paleoenvironments of Australia for the first time. We used this information, and phylogenetic-based analyses, to explore how changes in temperature and increasing aridity during the Neogene influenced the diversification of the Australian blindsnakes. We tested whether diversification rates differ between arid-adapted and mesic-adapted lineages.

**Taxon**: Typhlopidae, _Anilios_ blindsnakes

**Location**: Australia

**Materials and Methods**: We estimated the historical biogeography of blindsnakes using BioGeoBEARS. We synthesised multiple approaches to reconstruct paleotemperature and paleoaridity of Australia during the Neogene. We fit several birth-death models and estimate diversification rates under paleoenvironmental conditions using RPANDA. We further compare diversification rates between arid-adapted lineages versus mesic-adapted lineages using ClaDS and GeoHiSSE.

**Results**: Ancestral area estimation indicated Australian blindsnakes have tropical grassland origins. We found that Australia-specific regional paleotemperature and paleoaridity provided a better explanation for diversification rate variation than global paleotemperature. Specifically, our best-fitting model indicated that speciation rates of blindsnakes decreased with increasing aridity. We found no difference in diversification rates between arid- and mesic-adapted lineages.

**Main conclusions**: Soon after dispersing to Australia, the common ancestors of Australian blindsnakes diversified rapidly in mesic habitats during the early Miocene. However, as the continent became increasingly arid, diversification rates decreased. We found that shifts in the environment led to the emergence of two major clades: one remaining in primarily mesic habitats and the other adapting to the expanding arid biome. Our results emphasise the importance of both arid and tropical biomes as sources and sinks of diversification.

**Keywords**
BioGeoBEARS, diversification, GeoHiSSE, paleoaridity, paleoenvironment reconstruction, Scolecophidia

\newpage

# Introduction

The present-day distribution of extant species can provide clues about their biogeographic history and insight into the tempo of their diversification. In particular, one major aim of evolutionary biology is to understand the role that environmental factors have played in shaping rates of species diversification [@simpsonTempo1944]. A variety of environmental correlates have been associated with diversification processes across clades, including elevation [@esquerreHow2019;@boschmanMountain2022], CO<sub>2</sub> concentration [@kergoatOpposite2018], microhabitats [@bars-closelDiversification2017], temperature [@ricklefsGlobal2006;@mayhewBiodiversity2012;@condamineMacroevolutionary2013;@condamineClimate2019;@rollandContribution2019;@sunRecent2020], and aridification [@raboskyExceptional2007;@pepperGenetic2011;@deepakAridification2018;@catulloAridification2014]. However, despite advancements in our understanding of the evolutionary relationships and divergence times of many clades, our knowledge of the impact of paleoenvironmental change on these clades remains limited, particularly in regard to the regional scale at which many clades radiate [@eyresClimatic2021]. As such, it remains a key challenge to combine regional paleoenvironmental histories, with molecular phylogenies, to understand the drivers of diversification over deep-time scales.

<!-- Terrestrial ectotherms, such as squamate reptiles, are highly sensitive to changes in external environment and climate. This can influence their diversification rates through multiple mechanisms. For instance, warmer temperature may increase resource availability, which in turn increases ecological niche space [@clarkeClimate2006]. Spatial and temporal heterogeneity in temperature can lead to climatic niche divergence, such as local adaptations in different geographic areas [@pepperGeodiversity2013], and changes in activity times of prey and predators [@santosTemperature2011;@lawsonLatitudinal2014]. Change in climate, leading to glaciation or aridification, has been shown to cause fragmentation of continuous habitats leading to isolated populations, which may lead to species-level divergence [@oliverRocks2019]. Ectotherms have higher metabolic rates in warm climates (i.e. Kleiber’s law), which may increase mutation rates and ultimately greater speciation rates [@allenKinetic2006;@makarievaMean2008]. Finally, aridification may offer new ecological niche for species to exploit, particularly mesic-adapted lineages that have colonised from elsewhere. Despite many possible pathways, little is known about whether environmental change is associated with variable rates of diversification. -->

The Australian continent has one of the most diverse reptile faunas on the planet and this may have been driven by its rapid environmental change in the Neogene period. After a long period of relatively stable climate and mesic habitats, the aridification of continental Australia began around 20 Ma (mega annum) with the development of pockets of aridity [@martinCenozoic2006;@byrneDecline2011]. Subsequent extreme warming and cooling periods in the last ~5 Ma creating what is now known as the "outback" [@byrneBirth2008;@pepperLife2021]. During this period, as aridification progressed, arid-adapted vegetation expanded while mesic biomes became increasingly fragmented [@crispHow2013;@martinCenozoic2006]. This shift in the environment also led to a radiation of some reptile clades, such as elapid snakes and skinks, which diversified into morphologically and ecologically distinct lineages [@leeDiversification2016;@sandersMolecular2008;@raboskyExceptional2007]. Conversely, other reptile clades, such as blindsnakes and geckos, remain morphologically conserved and were thought to be species poor until recent molecular data suggested otherwise [@moritzCryptic2018;@zozayaAre2019;@marinHidden2013]. Furthermore, shifts in diversification rates due to changes in Australian temperature and aridity have been reported in some squamate reptile clades including skinks [@raboskyExceptional2007] and geckos [@fujitaDiversification2010;@oliverContrasting2014;@oliverIndependent2014;@pepperGenetic2011;@pepperPalaeoclimate2011] as well in frogs [@catulloAridification2014;@vidal-garciaConvergent2015], insects [@owenHow2017;@toussaintUnveiling2015], and plants [@crispHow2013;@kadereitMolecular2010;@millerComparative2013]. Yet, the role of aridification and temperature changes on diversification is highly variable, with some arid-adapted vertebrates radiating following aridification [@raboskyExceptional2007;@pepperPalaeoclimate2011;@brennanMass2017] while other mesic-adapted groups have had suppressed diversification or been driven extinct [@bryantLines2016;@cardilloPhylogeny2017;@owenHow2017]. Whether environmental change in Australia has had generally positive or negative influence on diversification rates in squamate reptiles is still unknown.

Today, the arid zone covers roughly 70% of the Australian landscape and provides habitat to 43% of all of Australia’s terrestrial reptiles [@byrneBirth2008].Of these, roughly a quarter of the snake species diversity in Australia is represented by blindsnakes of the genus *Anilios*, which currently comprises 47 described species  **47** described species [@uetzQuarter2021] (Tiatragul et al., submitted MS). All but three species of currenty described _Anilios_ are endemic to Australia, _A. torresianus_ and _A. leucoproctus_ also occur in southern parts of New Guinea, while _A. eryncinus_ has only been recorded in New Guinea and the Maluku Islands. 
Blindsnakes are morphologically conserved and previous phylogenetic and morphological studies have revealed that the true species diversity is underestimated [@marinHidden2013;@pyronSystematics2014;@hedgesTaxonomic2014;@ellisTyphlopid2019]. The diversification of _Anilios_ in Australia is reported to have originated from a single dispersal event by a mesic-adapted ancestor from Southeast Asia or Indonesia between 39 – 19 Ma [@vidalBlindsnake2010;@marinTracing2013]. 
However, phylogenetic analyses from Tiatragul et al. (2023) indicated that the closest relative to all *Anilios* is *Ramphotyphlops multilineatus*, a species that occur in Western New Guinea, Kai Islands, and Salawati, suggesting the Australian radiation may have a Melanesian origin
Despite representing a substantial portion of snake diversity, we know little about the macroevolutionary dynamics of blindsnakes. The monophyly of the genus makes _Anilios_ well suited to test hypotheses about how lineages diversified with dramatic environmental changes during the Cenozoic because they are geographically restricted and can be found in every biome on mainland Australia.

Advances in phylogenetic comparative methods and paleoenvironmental reconstructions allow us to better understand regional diversification dynamics. The influence of temperature on diversification rates has been empirically tested with birth-death models [@condamineMacroevolutionary2013;@kergoatOpposite2018;@condamineClimate2019;@rollandContribution2019]. Traditionally, these temperature data are reconstructed (interpolated) based on conversion of global concentration of oxygen isotopes [@condamineMacroevolutionary2013]. The relationship between diversification rates and aridity, however, have been attributed to some proxies such as the timing of aridification or global temperature rather than quantifying the extent of regional aridity itself [@dornStrange2014;@deepakAridification2018;@catulloAridification2014;@fujitaDiversification2010;@pepperGenetic2011;@pepperPalaeoclimate2011;@oliverIndependent2014;@oliverContrasting2014;@owenHow2017;@crispHow2013;@toussaintUnveiling2015].

To our knowledge no study has assessed aridity as a continuous variable in an environment-dependent birth-death model framework, probably due to the difficulty of quantifying paleoaridity. However, recent developments in spatially-explicit paleoenvironmental reconstructions from the lithological and fossil record as well as global circulation models (GCMs) now enables us to estimate paleotemperature and paleoaridity data specific to Australia in the time span of interest. The use of these data sets allows us to conduct empirical analyses of the impacts of regionalised environmental changes on diversification rates over time for the first time.

In this study, we aimed to infer the historical biogeography of the Australian blindsnake radiation and investigate the impact of paleoenvironmental changes, specifically changes in temperature and aridity associated with Neogene aridification, on diversification rates within this group. To do this, we fitted several birth-death models with reconstructed paleotemperature and aridity, and estimated diversification rates under these models. Additionally, we compared diversification rates between arid-adapted lineages versus mesic-adapted lineages to investigate if lineages that inhabit the arid biome have undergone faster or slower radiation.

#  Materials and methods

## Phylogenetic hypothesis and taxon sampling

We used the most updated phylogeny for *Anilios* and closely related outgroups from Tiatragul et al., submitted MS. The species tree was built with the shortcut coalescent method in ASTRAL III [@zhangASTRALIIIPolynomialTime2018] using maximum likelihood gene trees inferred in IQTREE from 4930 loci (322 Anchored Hybrid Enrichment, 4574 ultraconserved elements, and 34 legacy genes) for 35/47 recognised *Anilios* species. Tiatragul et al. (submitted) time-calibrated the phylogeny using 27 outgroup squamate fossils as primary calibrations using MCMCTree from the PAML package [@yangPAML2007] (see Tiatragul et al., submitted MS for details). Phylogenetic data in Tiatragul et al. (submitted) suggest the existence of at least three genetically distinct undescribed species, so we followed their designation and adjusted our taxon sampling fraction to 38/50 species. 
The genetically distinct OTUs that are not fully described species include three '_Anilios grypus_' (two in Western Australia and one in Queensland) and two '_Anilios ligatus_' (one from the eastern part of the range and one from the northern part) (Tiatragul et al., submitted MS). 
For ancestral biome estimation analyses, we included three outgroup blindsnakes species from Southeast Asia, Indonesia, and Melanesia, where the ancestor of *Anilios* has been hypothesised to have originated [@vidalBlindsnake2010]. The three species include *Ramphotyphlops multilineatus* (Western New Guinea, Kai Island, and Salawati), *Sundatyphlops polygrammicus* (Moyo Island, Indonesia), and *Acutotyphlops subocularis* (Bismarck Archipelago). We were not able to include a *Ramphotyphlops* from continental Southeast Asia in the time-calibrated phylogeny because insufficient data was available for divergence time estimation.

## Biogeography and biome transitions

To investigate the biogeographic history of blindsnakes, we used the R package 'BioGeoBEARS' v.1.1.2 [@matzkeBioGeoBEARS2013]. BioGeoBEARS provides a flexible, likelihood-based framework that models how branches in the phylogeny (representing extant taxa and their ancestors) have shifted among a set of discrete biomes over time. These shifts in biomes can occur within a single branch (anagenetic events, including processes such as dispersal and extinction) or at the time of branching (cladogenetic events involving sympatric, vicariant, and founder-event speciation). 
We divided Australo-Papua into nine discrete biomes and assigned each Australian species to one or more biomes based on their occurrence records from the Atlas of Living Australia database. Biome classifications were based on the World Wildlife Fund's (WWF) Terrestrial Ecoregions of the World data set [@olsonTerrestrial2001] and followed the nomenclature of published Australia-specific studies [@cardilloPhylogeny2017;@pavon-vazquezComprehensive2021].

Using the phylogeny and distribution data, we fitted models that vary in the kinds of cladogenetic and anagenetic events they model, including likelihood implementations of Dispersal-Vicariance Analysis (DIVALIKE) [@ronquistDispersalVicariance1997], Dispersal-Extinction-Cladogenesis (DEC) [@reeMaximum2008], and BayArea (BAYAREALIKE) [@landisBayesian2013] in BioGeoBEARS [@matzkeBioGeoBEARS2013]. We also fitted these three models, with two  additional relevant free parameters, both singly and together, for a total of 12 alternative models. The first additional parameter, j, assigns per-event weight to founder-events [@matzkeModel2014], and the second parameter, x, multiplies the probability of dispersal by distance to the power of x [@vandamEvaluating2016] (Table S1). We used likelihood ratio tests to determine if the additional free parameter was a justified improvement over the simpler base models and selected the model that best fitted the data based on corrected Akaike Information Criterion (AICc) scores. Based on the best model, we ran an additional 50 biogeographic stochastic maps to account for uncertainty in ancestral biome estimation and transition between biomes, and to estimate the frequency and direction of dispersal events across biomes [@matzkeBioGeoBEARS2013].

## Paleoenvironment reconstruction

We used two alternative approaches to estimate paleoclimate in Australia from 40 Ma to the present-day. The first approach used information derived from paleo-Köppen climate bands from @scoteseAtlas2021. Köppen bands are estimated from different fossil and lithological indicators and give a broad categorisation of spatial variation in major climate types (tropical, sub-tropical arid, warm temperate, cool temperate, and polar). We obtained estimated Köppen climate bands at 5 Ma intervals from 40 Ma to 0 Ma and used the sub-tropical arid Köppen band as a binary aridity index. We removed sites found in the ocean by using digital elevation models (DEMs) to mask sites below sea-level. To account for uncertainty in paleogeographic reconstructions of continental movements and sea-levels, we used two alternative DEMs [@straumeGlobal2020;@scotesePALEOMAP2018] to mask the ocean sites and provide two estimates of paleoaridity. To estimate paleotemperature from the Köppen bands, we followed the method of @scoteseAtlas2021 as implemented by @hagenEarth2021. Here, each Köppen climate band was assigned a temperature value based on present-day averages of those climate bands (tropical = 26, sub-tropical arid = 22, warm temperate = 16, cool temperate = 5, polar = -20). To account for global fluctuations in temperature, we then modified these values by a global temperature curve estimate from oxygen isotope data [@songSeawater2019]. Finally, to account for change of temperature with elevation, we used a lapse rate of 9.8°C/km. Following @hagenEarth2021, we model variation in the intensity of temperature oscillations during the Pleistocene. We also used the two alternative DEMs to give two alternative temperature curves using this approach [@straumeGlobal2020;@scotesePALEOMAP2018].

The second approach used a global circulation model (GCM) to obtain a model-based estimate of paleotemperature and paleoprecipitation [@valdesDeep2021]. The paleoclimatic data is based on a HadCM3L model which couples ocean, atmosphere and vegetation and simulated at 5 Ma intervals based on the paleogeography of @scoteseAtlas2021. To make it comparable with the binary aridity index derived from the Köppen band approach, we converted the paleoprecipitation into an aridity index by thresholding the precipitation data. We applied a threshold of 500mm/year, below which is considered arid, above which was considered non-arid [@holzapfelDeserts2008;@waltonArid1969]. Each of the three global paleoaridity and paleotemperature maps, from the two approaches, were projected into a Behrmann equal-area projection using a spatial resolution of 110 km x 110 km. We then temporally interpolated the data from 5 Ma intervals to 100 ka using bilinear interpolation. Finally, we subset the data to include only the continent of Sahul from 40 Ma to the present-day. From these maps, we obtain multiple Sahulian curves for paleotemperature by taking the mean, minimum and maximum values within Sahul at each timestep. For paleoaridity, we took the mean value, as minimum and maximum showed little variation in the binary metric.

<!--- Because reptiles are sensitive to extremely low and high temperatures, we decided to summarise the reconstructed paleo temperature data into three different data sets: minimum temperature, maximum temperature, and mean temperature. --->

## Time- and environment-dependent diversification analyses

To evaluate how diversification rates correlated with climate and aridity through time, we fitted several time- and environment-dependent likelihood diversification models using the fit_bd and fit_env functions in the R package 'RPANDA' v.2.0  [@morlonRPANDA2016] to the _Anilios_ phylogeny. Specifically, we fitted different combinations of pure-birth and birth-death models with speciation and/or extinction rates being constant or otherwise varying linearly or exponentially with paleoenvironment as described above. We fitted 12 diversification models based on reconstructed paleoenvironmental data to the phylogeny including three paleotemperature (mean, minimum, and maximum) and the paleoaridity index from each paleoenvironmental reconstruction method. We also fitted an additional environmental dependent diversification model with the widely use mean global paleotemperature data set [@condamineMacroevolutionary2013] to provide comparison to our reconstructed paleoenvironmental data. To account for missing taxa, we used the fractioning function (f = 38/50) in RPANDA. We then assessed which diversification model best fit the data based on AICc scores.

## Geographic range and rates of diversification

To test whether diversification rates differ between arid-adapted and mesic-adapted lineages, we used two approaches. First, we estimated speciation rates at the origin of each branch (henceforth branch-specific rates) using ClaDS function in the Julia package 'PANDA' v.1.6.1 [@malietFast2022]. ClaDS is designed to capture frequent shifts in speciation rates with small effects. At each speciation event, daughter lineages are assigned its own speciation rate sampled from a distribution parameterised on the parental rate [@malietFast2022].

In the second approach, we fitted four state-dependent speciation and extinction (SSE) models to the phylogeny and geographic range data under the GeoSSE [@goldbergPhylogenetic2011] and GeoHiSSE [@caetanoHidden2018] framework using the R package 'hisse' v.2.1.9 [@beaulieuDetecting2016]. 
Since these SSE models only allow two states, we restricted our geographic states to arid and mesic. Based on the same categories we used in our BioGeoBEARS analyses, we classified species that primarily occur in the arid biome as arid (State 0; n = 8), species that occur primarily in other biomes as mesic (State 1; n = 25), and those that occupy both the arid and mesic biomes as widespread (State 01; n = 5).

The four SSE models differed in whether they allowed diversification rates to be dependent on geographic states (range-dependent) or independent of them (range-independent), and whether they allowed an unmeasured trait ('hidden state') to influence diversification rates [@beaulieuDetecting2016;@caetanoHidden2018]. 
Model 1 assumed equal diversification rates regardless of geographic state. Model 2 allowed geographic range to affect diversification rates.
Models 3 and 4 were based on models 1 and 2, but they also allowed  a hidden state to contribute to diversification rates. By incorporating hidden states, we could assess whether geographic range or an unmeasured trait contributed more to the observed diversification. For each model, we assumed uniform taxon sampling for 89% of arid lineages, 71% of mesic lineages, and 83% of widespread lineages, based on the number of species in the phylogeny and all currently described species of _Anilios_. We then calculated AIC to identify the best-fitting models. Following the approach in @caetanoHidden2018, we averaged the estimates of diversification parameters across all models according to their AIC weights.

PANDA was implemented in Julia v.1.8.1 [@bezansonJulia2012], all other analyses were performed in R v.4.2.1 [@rcoreteamLanguage2022]. Figures were prepared using 'ggplot2' v.3.4.0 [@wickhamGgplot22016]. All map graphics were created using the Free and Open Source QGIS v.3.6. We arranged and finalised all figures using Adobe Illustrator v2022.

<!-- Anilios_bituberculatus - distributed in the arid zone but also many occurrence in coastal areas in southern Australia. Classified as mesic.
Anilios_diversus - mostly in Top End with some distribution in the interior. Classified as mesic
Anilios_grypusET - occur in eastern edge of the arid biome. Also occur in tropical savannah. Classified as mesic.
Anilios_grypusNW - closest to the range of A. grypus in the tropical grassland but also occur in Pilbara. Classified as arid. 
Anilios_hamatus - species description says widespread through 'southern arid and semi arid regions'. Classified as Arid.   -->

# Results

## Ancestral biome reconstruction and transitions

<!-- ## values and info from 04_BioGeoBEARS_bsm_plots.R -->
Our BioGeoBEARS analyses all inferred tropical grassland as the ancestral biome of *Anilios*, but the ancestral biome for the common ancestor between *Ramphotyphlops multilineatus* and *Anilios* is ambiguous (Fig. \@ref(fig:fig-biogeobears)a and Fig. S1). We found that the DEC+j+x model was the best-fitting biogeographical history model (Δ AICc = -4.47, Table S2). A likelihood ratio test (LRT) showed that both founder events (+j parameter) and distance between biomes (+x parameter) together provide statistically significant improvements over the simpler DEC, DEC+x, and DEC+j models (χ^2^, DF = 2, D-statistic = 26.7, p-value < 0.001).

Based on 50 BioGeoBEARS biogeographic stochastic mappings (BSM) of the DEC+j+x model, we estimated that an average of 71.2 ± 1.3 (mean ± SD) total biome shifts occurred throughout the phylogeny (see Table S3). Of these 40 ± 0 were cladogenetic (branching) events compared to 31.2 ± 1.3 anagenetic (within a single branch) dispersal events. Out of the 40 branching events, in 23.8 ± 2.1 events sister species maintained their ancestral biome (“narrow sympatry”) and in 5.2 ± 2.3 events one sister species maintain their ancestral biome while the other sister occupies only a subset of the ancestral range (“subset sympatry”). The remaining branching events were made up by 6.1 ± 1.5 founder events where one of the sister species occupies a new range, and 4.9 ± 1.3 vicariance events where the sister species split their ancestral range. Our model also indicate that range expansion is expected to decrease as distance increase (free parameter x = -1.3).

Based on current distributions from ALA, most Anilios species are found entirely or partially in the tropical grassland biome (n = 20) followed by the arid biome (n = 13) and Mediterranean biome (n = 10). Apart from A. silvia, A. leucoproctus, and A. proximus that appeared to form independent divergent lineages (Fig. \@ref(fig:fig-biogeobears)a), we identified divergent biogeographic transitions between two major groups of species, which we refer to as Group 1 (n = 21 lineages) and Group 2 (n = 14 lineages). The common ancestor of Group 1 remained mostly in the tropical grassland biome, then 12/21 lineages expanded or transitioned into other biomes. The common ancestor of Group 2 first transitioned into the arid biome, four lineages remained in the arid biome while ten lineages transitioned or expanded into other biomes.

The tropical grassland biome is the largest source of outward dispersal (16.2 ± 4.9 events; 46.6%) followed by the arid biome (9.7 ± 3.2 events; 26%)(Fig. 1b). On the other hand, the largest sinks for all inward dispersals are the tropical forest biome (7.9 ± 4.2 events; 21.2%) and the Mediterranean biome (7.0 ± 2.9 events; 18.8%). The greatest directional transition is from the arid biome to the Mediterranean biome (5.3 ± 1.5 events) and from the tropical grassland biome into the temperate forest biome (5.1 ± 1.3 events) (Fig. \@ref(fig:fig-biogeobears)c).

```{r fig-biogeobears, fig.cap= 'Ancestral biome estimation and transitions of Australo-Papuan blindsnakes in BioGeoBEARS, using the DEC+j+x model. (a) Pie charts based on 50 biogeographic stochastic map (BSM) representing probabilities of each biome at speciation event. Inset map (left) shows the current distribution of biomes. Inset maps to the right of the phylogeny shows distribution of species in Group 1 and Group 2. (b) Bar plot showing frequency and composition of outward transitions and (c) inward transitions from Australian biomes and combined non-Australian islands. Abbreviations: TrG = Tropical Grassland, TmF = Temperate Forest, AR = Arid, MD = Mediterranean, TrF = Tropical Forest, and TmG = Temperate Grassland.'}

knitr::include_graphics(paste0(WD, '/assets/figures/biogeobears.jpg'))
```

## Paleoenvironment estimates

Paleotemperature estimates for Australia differed between the three different methods (Fig. \@ref(fig:fig-paleoclimate-compare)a). However, all three paleotemperature models indicate Australia was hotter than the average global temperature (Fig. \@ref(fig:fig-paleoclimate-compare)a; Fig. S2). Paleotemperature in Australia were also relatively constant compared to global temperature until the Plio-Pleistocene, a period characterised by fluctuations corresponding to the glacial cycles. Reconstructions of paleoaridity differed slightly between the two Köppen band methods, but these differed substantially to the aridity index derived from the GCM data (Fig. \@ref(fig:fig-paleoclimate-compare)b). Both Köppen band-based aridity indices show increasing aridity from 10 Ma.

```{r fig-paleoclimate-compare, fig.cap= 'Paleoclimate estimation based on three different methods and RPANDA. (a) Estimated mean temperature over time in Australia compared to RPANDA global paleotemperature data. (b) Estimated mean aridity index over time based on precipitation threshold. Abbreviations Pli = Pliocene and Ple = Pleistocene.'}

knitr::include_graphics(paste0(WD, '/assets/figures/paleo_data_compare.jpg'))

```

## Environment-dependent diversification

Among all the birth-death, time-dependent, and environment-dependent models fitted in RPANDA, we found that environment-dependent models were preferred. Of the fitted environment-dependent models, we found the exponential environment-dependent speciation and constant extinction (BEnvDCST) model based on paleoaridity estimated from @scotesePALEOMAP2018 provided the best fit to the phylogeny (AICc = 238.43; see Table 1). This best-fitting model indicates that an increase in aridity is exponentially associated with a decrease in speciation rate (λ = 1.32, α = -0.14) and a constant background extinction rate (µ = 0.23) (Fig. \@ref(fig:fig-RPANDA)). The second and third most competitive models (ΔAICc < 2) also found that an increase in aridity is exponentially associated with a decrease in speciation rate, but without estimating extinction (see Table 1). Among the exponential environment-dependent models, we found that Australia-specific paleotemperature data provided a better fit to our phylogeny than global paleotemperature data (Table S4).

<!-- {r Table-RPANDA-TOP10} -->

**Table 1**. Results of the 10 most supported RPANDA diversification models on the Australian blindsnake radiation. The table includes the models (BEnvVar = Speciation rate varying with environment; DCST = Constant extinction rate; DEnvVar = Extinction rate varying with environment; (EXPO) = rates varying exponentially with environment; (LIN) = rates varying linearly with environment) number of parameters, the estimated log-likelihood (logL), the corrected Akaike information criterion (AICc), and corresponding parameter estimates (λ = speciation rate at present, α = parameter controlling the dependency of speciation rate on time or paleoclimate, µ = extinction rate at present, ß = parameter controlling the dependency of extinction rate on time or paleoclimate). Sources of paleoenvironment presented in the table: ^1^ @scotesePALEOMAP2018  and ^2^ @straumeGlobal2020

| TOP10       | Description |
| ----------- | ----------- |
| Header      | Title       |
| Paragraph   | Text        |

```{r fig-RPANDA, fig.cap= 'Effect of aridity change on the diversification of Anilios estimated from the phylogeny. (a) The aridity index through time calculated from reconstructed paleoclimate (dotted line), and lineage through time estimated from the phylogeny (solid line). (b) Fitted diversification rates through time reconstructed from parameter estimates from the best diversification model (speciation rates vary exponentially with paleoaridity and constant extinction rate; “BEnvVar + DCST (EXPO)”). (c) Fitted diversification rates at different aridity index values calculated from parameter estimates of the previous model. Increased aridity is associated with decrease net diversification rates in this model..'}

knitr::include_graphics(paste0(WD, '/assets/figures/RPANDA_fit_arid.jpg'))
```

## Diversification rates between arid and mesic biomes

We applied ClaDS and SSE models to evaluate whether diversification rates differ between lineages in arid, mesic, and widespread ranges (Fig. \@ref(fig:fig-clads-sse)). 
Results from our ClaDS analysis show branch-specific diversification rates was high when the genus first diversified and gradually decreased towards the present (Fig. \@ref(fig:fig-clads-sse)b). ClaDS also indicate mean branch-specific rates did not differ significantly between the geographic ranges (Fig. \@ref(fig:fig-clads-sse)c).

In our SSE analyses, we found that the range-independent model fit the data better than models where diversification rates were dependent on geographic range (Table 2). The best-fitting geographic range-independent model included a hidden unobserved state, indicating that diversification rates are better explained by this hidden state. Compared to the three other models, the best-fitting model accounted for 95% of the AIC weight and had an AIC value 7.02 units lower than that of the second best-fitting model (Table 2). The weighted average parameter estimates from all models indicate that both speciation rates and turnover rates did not differ between arid and mesic ranges (Fig. S3 and S4; Table S5).

<!-- Table-geohisse -->

**Table 2**. Diversification rate models investigated in the GeoSSE and GeoHiSSE framework. AIC scores and AIC weights indicate which models best fit the data.

| GeoHiSSE      | Description |
| ----------- | ----------- |
| Header      | Title       |
| Paragraph   | Text        |

```{r fig-clads-sse, fig.cap= 'Diversification rates estimates from ClaDS and SSE models. (a) Branch-specific rates (log(λi)) across the phylogeny. (b) Mean branch-specific rates over time. Thin grey lines: individual draws from the MCMC. Blue lines: 95% credibility interval in the MCMC. Dotted green line: inferred mean rate through time estimated from the MCMC iterations. (c) Density plot of branch-specific rates by geographic range from ClaDS analysis.'}

knitr::include_graphics(paste0(WD, '/assets/figures/fig-clads-sse.jpg'))
```

# Discussion

With 47 recognised species, the mostly Australian-endemic blindsnake genus *Anilios* is currently the most diverse genus of Scolecophidia in the world. In this study we estimated that the ancestral species in Australia occupied the tropical grassland biome, most likely as a result of invasion from other tropical mesic areas including New Guinea, Bismarck Archipelago, and the Lesser Sundas. Within a few million years, this ancestral lineage split with one group staying in the tropical grassland biome while the other group transitioned into the arid biome. Subsequent biome transitions in these two groups resulted in the distribution of biomes we observed across the phylogeny today. We also found that the paleoaridity index provides the best fit to the diversification dynamics observed in our time-calibrated phylogeny. Although aridification has been linked to elevated diversification rates in other plants and animals [@brennanMass2017;@raboskyAnalysis2014;@hernandez-hernandezAridification2014;@jhaPhylogeny2021], our results suggest that aridification decreased diversification rates for Australian *Anilios* blindsnakes.

## Historical biogeography of *Anilios*

Our result supports the hypothesis that the ancestor of the Australian clade first dispersed to the mesic part of the Australian continent, which is consistent with previous studies [@marinTracing2013;@vidalBlindsnake2010]. The observed pattern where the ancestral lineage first inhabited tropical-mesic regions before later expanding into other biomes as they emerge has also been documented in other radiations of Australian squamates [@chappleParallel2004;@melvilleEvolutionary2011;@oliverPhylogenetic2017]. 
Many Australian groups originated in Eurasia and crossed into Australia after the continental plates collided and caused the uplift of a number of 'stepping stones' including the Sunda islands, New Guinea, and the Soloman islands. Our historical biogeographic estimation of *Anilios* is consistent with the hypothesis that this group arrived in Australia during the Miocene, first occupying tropical grasslands of Northern Australia before diversifying throughout the continent. However, whether the Australian radiation was a result of an oceanic dispersal from Southeast Asia or Indonesia, as proposed by @vidalBlindsnake2010, is equivocal. Our results indicate the common ancestor of Australian _Anilios_, _S. polygrammicus_, _R. multilineatus_ and _A. subocularis_ most likely occurred either in the Bismarck Archipelago or New Guinea. Increasing taxonomic sampling, particularly by including the New Guinean endemic _Anilios erycinus_ and more species from *Ramphotyphlops* and *Acutotyphlops* will enable us to make more informed inferences about the origin of the group prior to colonising Australia.

Tiatragul et al., (submitted) proposed that the earliest divergence of *Anilios* occurred around 25 Ma (28 – 22 Ma; 95% HPD), which is slightly older than 21.9 Ma (30.2 – 15.1 Ma; 95% Bayesian credibility interval) proposed by previous studies [@marinTracing2013;@vidalBlindsnake2010]. Our BioGeoBEARS analyses suggest that _Anilios_ gradually diversified in the tropical grassland (Group 1) and arid biome (Group 2) during a time when Australia was relatively wet and humid [@byrneBirth2008]. This indicates that many species found in arid biomes, particularly in Group 2, diversified in the arid biome prior to the onset of widespread aridification in the late Miocene [@byrneBirth2008].

We hypothesise that the contemporary distribution of _Anilios_ and other blindsnakes in arid biomes is explained by pre-adaptation to living in such environments. blindsnakes have been found in xeric areas around the world, such as _Rena humilis_ in the desert of North America and _Myriopholis macrorhyncha_ in semi-arid areas from east Africa to the Middle East [@sindacoReptiles2013]. These arid-adapted blindsnake species share specialisations in diets and morphology with Australian _Anilios_, including a diet of pupae and larvae of social insects [@webbDietary1993;@cundallFeeding2000], smooth scales that protect them from biting adult ants/termites, relatively small body size compared to other snakes, and fossorial lifestyle. A combination of these traits allows blindsnakes to avoid direct heat exposure and risk desiccation  in harsh arid environments.

## Regional vs global paleoenvironment 
Our reconstructed Australia-specific paleotemperature data provided a better fit to the environment-dependent diversification models we considered in RPANDA than the global temperature data set [@condamineMacroevolutionary2013], although models that used paleoaridity provided better fit (Table 1; Table S4). In the context of our birth-death models, the absolute value of temperature might not have made much of a difference because our environment-dependent models fit the spline rather than absolute temperature. However, models that take in account thermal niche tolerance, such as NEMo [@huaAdapting2022], will benefit from paleoclimate reconstructions such as these to inform limits (boundaries). As it stands, current estimates of climate at deep timescales are still developing and we see strong differences in climate estimates from different approaches. For example, while both GCM-based and Köppen band-based approaches showed a gradual increase in temperature from 25 – 5 Ma before a decrease from the Pliocene to the present-day, the GCM-based approach showed a much sharper increase in temperature (Fig. \@ref(fig:fig-paleoclimate-compare)a). Similarly, aridity estimates from the GCM-based approach were starkly different to the Köppen band approaches (Fig. \@ref(fig:fig-paleoclimate-compare)b). This may be because deep time GCM models are currently aimed at estimating general global fluctuations in climate, rather than regionally specific ones, which may be poorly approximated. Instead, the Köppen band approach is data driven and is based on known indicators of aridity from around the world [@boucotPhanerozoic2013]. This could explain why the Köppen band approach for aridity provided the strongest explanation for the diversification rate data. Because climate estimates can vary widely, we suggest future studies that aim to understand how aridification affect diversification rates in Australia consider using these (and other) paleoclimate models to inform model inferences.

## Aridification and diversification rates

Our best-fitting diversification model show that speciation is negatively correlated with aridity (Fig. \@ref(fig:fig-RPANDA)c). Our reconstructed paleoaridity index data indicate that aridity increased from ~25 Ma following the opening of the Southern Ocean and the onset of the Antartic circumpolar current [@byrneBirth2008]. A particularly sharp increase from 10 Ma followed from the "Hill Gap" period of cooling [@hillHistory1994], and intensification from 5 Ma onwards following the Pliocene transition to a contemporary arid Australian climate [@byrneBirth2008]. Out of the 38 lineages we considered, 5/35 divergence events occurred in the last 10 Ma when the continent intensively aridified. In contrast, 31/35 divergence event occurred between 25 – 10 Ma when the continent was getting warmer but not entirely arid. Indeed 12/16 lineages that have current distributions in the arid biome diverged before 10 Ma. This pattern is consistent with other studies on vertebrates that have shown that the divergence time for some arid-adapted taxa pre-dated the intense aridification in the last 5 Ma [@byrneBirth2008;@catulloAridification2014].

The observed decrease in diversification rate with increasing aridification may be linked to decreasing speciation rates or increasing extinction rates. One hypothesis is that mesic-adapted ancestral blindsnakes may have been widespread in the past, but populations later became extinct due to extreme aridity in the last 5 Ma [@fujiokaAustralian2009;@byrneDecline2011].
An extensive population genetics study of blindsnake by @marinHidden2013 proposed that during widespread aridification, mesic-adapted lineages went extinct or were confined to mesic refugia, such as the Pilbara and Kimberley, where they became pre-adapted to seasonal aridity, which enable subsequent invasion into the truly arid environments.
Additionally, long naked branches in our current phylogeny, particularly in Group 1, suggests past extinction events or intraspecific diversity not yet recognised.
Our best-fitting models of diversification support constant extinction rates with decreasing rates of speciation. 
However, the lack of *Anilios* fossils makes extinction rates difficult to estimate [@loucaWhy2021;@raboskyExtinction2010]. An alternative explanation for 
the negative relationship between aridification and diversification is decreasing speciation rates through time. This could be driven by the indirect effect of aridification on population size [@goldieEvolutionary2010]. More arid sites have lower primary productivity and may support smaller population sizes. Under the integrated evolutionary speed hypothesis [@gillmanPatterns2013;@rohdeLatitudinal1992], smaller population sizes are predicted to have slower rates of molecular evolution and subsequently lower rates of speciation. Lower rates of molecular evolution have been found in Australian arid zone plants compared to mesic zone plants [@goldieEvolutionary2010]. 
Yet, here we found no difference between arid zone and mesic zone speciation rates in blindsnakes which is inconsistent with this hypothesis. 
Finally, another potential explanation for the negative correlation between diversification and aridity is the limited sampling of Australian blindsnakes, particularly in major sand deserts [@pepperLife2021] or even in areas of high cryptic diversity such as the Kimberley [@ellisTyphlopid2019]. The current phylogeny is missing 12 described species (Tiatragul et al., submitted MS), such as the arid-adapted _A. fossor_ and _A. minimus_, and the true diversity has been estimated to exceed 56 species [@marinHidden2013;@marinTracing2013;@ellisTyphlopid2019]. It is possible that unrecognised species may be hidden under nominal species and that these missing lineages have diversified more recently when aridity is high. If all suspected cryptic OTUs were treated as recognised species, we could see a different relationship with aridification.

## Diversification rates in arid vs mesic biomes
Several studies have shown that arid-adapted species originated either from a single ancestor present in Australia before the Miocene, or through multiple independent expansion into the arid zone [@byrneBirth2008]. Our ancestral biome analyses show that several mesic ancestors of _Anilios_ blindsnakes have diverged into the arid zone evident by multiple arid species throughout the phylogeny. This pattern is consistent with several other groups of reptiles that started to diversify in the Miocene including monitor lizards [@brennanPhylogenomics2021;@pavon-vazquezCompetition2022], _Sphenomorphus_ and _Eugongylus_ skinks [@skinnerLygosomine2011], elapid snakes [@sandersMolecular2008], pythons [@esquerrePhylogenomics2020;@esquerreEvolutionary2022] and several geckos [@ashmanDiversification2018;@fujitaDiversification2010;@oliverIndependent2014]. For these terrestrial reptiles, increased rates of speciation for arid-adapted taxa have been attributed to species filling niches in the saxicolous and xeric habitats that emerged with the contraction of mesic forest throughout the Miocene [@sandersMolecular2008;@hugallPhylogeny2008]. However, results from our weight-averaged SSE analyses on the Australian blindsnake phylogeny indicated that speciation, extinction, and turnover rates did not differ between lineages in arid and mesic biomes. This lack of difference between biomes could be due in part to unrecognised species from the often under sampled arid zone [@pepperLife2021]

<!-- A potential explanation is that fossorial species may not be subjected to the same selective pressures present in the arid zone, such as extreme temperatures [@pepperLife2021], as much as surface dwelling species do. For instance, because blindsnakes live in subterranean microhabitat with relatively stable climate compared to the surface [@kinlawReview1999;@nevoAdaptive1979], modification from the already specialised phenotype may not be selected for [@cyriacDigging2018]. Future studies that investigate phenotypic variations across all blindsnakes will further our understanding of how these snakes evolved with seemingly conserved phenotypes. -->

<!---
- Future phylogeographic study should consider densely sampling species to see whether there are more expansion into or out of arid zone for widespread taxa.
Warmer and dryer environment may not directly influence speciation rate because blindsnakes are physically adapting to warmer temperatures.
However, climatic fluctuations can contract and fragment available niche for blindsnakes and thus influence diversification rates.
- Microhabitat better explains diversification than climate, but typhlopidae has the second highest speciation rates among fossorial snakes [@bars-closelDiversification2017].
- Interestingly blindsnakes have diversified greatly compared to other fossorial vertebrates.
  - It is also possible that fossorial species are more cryptic and thus the full extent of their diversity is under appreciated.

##### Geographical things
- Geographic barrier for mesic adapted lineages in the temperate and tropical forests may have difficulty crossing the lines [@bryantLines2016]
- Certain regions with complex topographic features tend to harbour more diveristy than flat featureless desert areas [@pepperPalaeoclimate2011].
- Lineages that have evolved more recently in the arid zone became extinct with the Plio-Pleistocene glacial cycles leaving only species that have survived in pockets of refugia.

- The early burst of speciation is observed in lineages that colonise new areas. For example ...
- Extinction is a big cavaet especially because we suspect there are more species. Arid adapted lineages may have experienced higher extinction OR not well sampled in remote Western Australia. Most species are sampled next to roads [CITE Xia's paper or do supplemental analyses].
--->
<!---
@pepperGenetic2011 - mountain ranges harbour more diversity than deserts. arid zone vertebrates show little genetic diff. But recent population expansion in the desert. 
@catulloAridification2014 - multiple independent origin from mesic or monsoonal ancestors into the arid zone. Lineages older than intense aridification. 
@fujitaDiversification2010 - independent origins of arid zone lineages in _Heteronotia_
@oliverIndependent2014 - multiple independent invasion into the arid zone.
@owenHow2017 - cicada increase speciation in arid zone compared to mesic zone (SSE)
@toussaintUnveiling2015 - diving beetle boom and bust diversification. Invaded arid zone and changed ecologically
@deepakAridification2018 - India accelerated speciation with aridification from the late Miocene. Probably due to emergence of open habitats
--->

## Limitations
Although we did not detect conflicting signal between our analyses, some of our results, particularly those that estimate diversification rates (GeoHiSSE, ClaDS, and RPANDA), may be biased due to the small sample size of our _Anilios_ phylogeny [@caetanoHidden2018;@malietModel2019;@sunEstimating2020]. However, this near-complete phylogeny of _Anilios_ is the most comprehensive data set available at this time. 
We suspect there remains several undescribed species of blindsnakes in Australia and New Guinea that may belong in the genus _Anilios_. Getting genetic data from _A. erycinus_ will allow us to place the species among other _Anilios_ and provide a more comprehensive picture of the evolution of this genus.

Our BioGeoBEARS ancestral biome estimation is limited by the areas (e.g. in this case biomes) to which we can assign our extant species. By following the WWF Bioregions maps, we had to assign species that occur in extremely different habitats in the same biome. For example, *A. longissimus*, a species that has only been collected from 30 m underground on Barrow Island located 50 km the northwest off the Pilbara coast of Western Australia, shares the same biome with *A. centralis* from the arid interior of the continent. Current methods do not allow us to account for heterogeneity of the biome even though we know the arid biome consists of everything from flat, seemingly featureless, desert in the interior to topologically complex mountain ranges on the fringes [@pepperLife2021;@pepperGenetic2011]. However, modern software under development such as PhyBEARS may be able to account for more complex geological processes such as the movement of continents. Finally, (mass) extinction likely played important roles in the pattern of diversity we observed today [@oliverPhylogenetic2017]. Out of all our comparative analyses, only the SSE and some RPANDA models estimated extinction rates.

<!---
- Without extensive sampling we can not confidently trace the pathways in which _Anilios_ colonised Australia. Due to the coarseness of our biomes, we also cannot infer exactly where the ancestral species landed in Australia.

Therefore, we cannot make any conclusions about transitions between New Guinea and Australia even though at least two species (i.e. _A. torresianus_ and _A. leucoproctus_) are found between southern New Guinea and Far North Queensland 
--->

# Conclusion

As we continue to generate more genetic data to build ever more extensive phylogenies, there is a growing need to develop methods, and complimentary paleoenvironmental data sets, that allow us to test hypotheses about how different macroevolutionary processes influence diversification. We inferred that tropical grassland is the ancestral biome of Australian _Anilios_ and acted as a major source of biodiversity to other biomes. Compared to other Australian reptile radiations in the Miocene, we observe a burst of diversification when the lineage first arrived on the continent and gradual decrease in diversification as the continent aridified. 
We showed that regionalised paleoaridity data provides a better fit to our Australian blindsnake phylogeny than global paleotemperature, highlighting the need to include and improve estimates of continental scale paleoenvironmental data in diversification analyses. To our knowledge, this is the first study that empirically fit diversification models to Australia-specific reconstructed paleoaridity data. We believe these regional paleoclimate models will be of interest for studies where species are geographically restricted.

\newpage

# Data availability statement

Data and analyses scripts are available from the Dryad Repository (https://doi.org/10.5061/dryad.t4b8gtj5x)

# Figure legends

Fig. Ancestral biome estimation and transitions of Australo-Papuan blindsnakes in BioGeoBEARS, using the DEC+j+x model. (a) Inset map (left) shows biome classifications of the regions considered in this study. Coloured boxes at the tips correspond to biomes on the map. Pie charts represent ancestral biome reconstruction based on 50 biogeographic stochastic maps (BSM) for each given nodes. Colours present on the pie charts but not in the biome map indicate mixture of more than one biome. Coloured boxes below pie charts indicate the most probable biome(s) at visually ambiguous nodes. Inset maps (right of the phylogeny) shows distribution of species in Group 1 and Group 2. (b) Barplots showing frequency and composition of outward transitions and (c) inward transitions from Australian biomes and combined non-Australian islands. Abbreviations: TrG = Tropical Grassland, TmF = Temperate Forest, AR = Arid, MD = Mediterranean, TrF = Tropical Forest, and TmG = Temperate Grassland.

Fig. Paleoclimate estimation based different data sets in gen3sis and RPANDA. (a) Estimated mean paleo temperature in Australia compared to RPANDA global paleotemperature data. (b) Estimated mean aridity index based on precipitation threshold. Abbreviations Pli = Paleocene and Ple = Pleistocene.

Fig. Effect of aridity change on the diversification of *Anilios* estimated from the phylogeny. (a) The aridity index through time calculated from reconstructed paleoclimate (dotted line), and lineage through time estimated from the phylogeny (solid line). (b) Fitted diversification rates through time reconstructed from parameter estimates from the best diversification model (speciation rates vary exponentially with paleoaridity and constant extinction rate; “BEnvVar + DCST (EXPO)”). (c) Fitted diversification rates at different aridity index values calculated from parameter estimates of the previous model. Increased aridity is associated with decrease net diversification rates in this model.

Fig. Diversification rates estimates from ClaDS and SSE models. (a) Branch-specific rates (log(λi)) across the phylogeny. (b) Mean branch-specific rates over time. Thin grey lines: individual draws from the MCMC. Blue lines: 95% credibility interval in the MCMC. Dotted green line: inferred mean rate through time estimated from the MCMC iterations. (c) Density plot of branch-specific rates by geographic range from ClaDS analysis.

# Tables.

Table 1. Results of the 10 most supported RPANDA diversification models on the Australian blindsnake radiation. The table includes the model, number of parameters, the estimated log-likelihood (logL), the corrected Akaike information criterion (AICc), and corresponding parameter estimates (λ = speciation rate at present, α = parameter controlling the dependency of speciation rate on time or paleoclimate, µ = extinction rate at present, ß = parameter controlling the dependency of extinction rate on time or paleoclimate).

Table 2. Diversification rate models investigated in the GeoSSE and GeoHiSSE framework. AIC scores and AIC weights indicate which models best fit the data.

\newpage

# Biosketch

Sarin Tiatragul is broadly interested in learning about how organisms came to be and their interaction with the environment. This work is part of his PhD work at the Australian National University on the macroevolution of Australian blindsnakes.

Alexander Skeels is a postdoctoral researcher who works on the evolution of biodiversity and how deep-time processes such as plate tectonics and climate change shape diversification patterns in plants and animals.

J. Scott Keogh is a professor of evolutionary biology with wide ranging interests in molecular phylogenetics and molecular ecology.

# Author contributions: 
S.T and J.S.K conceived the ideas; S.T and A.S collected the data; S.T analysed the data; S.T led the writing; A.S and J.S.K edited the manuscript.

# References
<!-- # References -->